Msprime 1.0: an efficient and extensible coalescent simulation framework

### 2021-07-23 paper.tex
added diagram of mutated trees

\newcommand{\jkcomment}[1]{\textcolor{red}{#1}}

@@

% Explain the background: what have people done classically?
Because coalescent simulations are usually concerned with
neutral evolution (see the \nameref{sec-selection} section, however)
the problem of generating synthetic genetic variation can be decomposed into
two independent steps:
simulating genetic ancestry (the trees), and then simulating
variation by superimposing mutation processes on those trees.
Indeed, to call a mutation ``neutral'' means that its presence or absence
does not affect demography.
A number of programs exist to place mutations on trees:
in addition to \phastsim~\citep{demaio2021phastsim},
the classical \texttt{Seq-Gen} program~\citep{rambaut1997seq}
allows a range of different models of sequence evolution to be imposed on a given
tree in Newick format, and there are various
extensions to the basic nucleotide
models~\citep[e.g.][]{cartwright2005dna,fletcher2009indelible}.
Population genetic simulators have tended to
include their own implementations of mutation simulation, with
most just supporting the infinite sites
model~\citep[e.g.][]{hudson2002generating}
but with several supporting a range of different models
of sequence
evolution~\citep[e.g.][]{mailund2005coasim,excoffier2011fastsimcoal,
virgoulay2021gspace}.

>>>>

\newcommand{\jkcomment}[1]{\textcolor{red}{#1}}
\newcommand{\plrcomment}[1]{\textcolor{purple}{#1}}

@@

% Explain the background: what have people done classically?
Because coalescent simulations are usually concerned with
neutral evolution (see the \nameref{sec-selection} section, however)
the problem of generating synthetic genetic variation can be decomposed into
two independent steps:
simulating genetic ancestry (the trees), and then simulating
variation by superimposing mutation processes on those trees
(see figure~\ref{fig-mutated-trees}).
Indeed, to call a mutation ``neutral'' means that its presence or absence
does not affect demography.
A number of programs exist to place mutations on trees:
in addition to \phastsim~\citep{demaio2021phastsim},
the classical \texttt{Seq-Gen} program~\citep{rambaut1997seq}
allows a range of different models of sequence evolution to be imposed on a given
tree in Newick format, and there are various
extensions to the basic nucleotide
models~\citep[e.g.][]{cartwright2005dna,fletcher2009indelible}.
Population genetic simulators have tended to
include their own implementations of mutation simulation, with
most just supporting the infinite sites
model~\citep[e.g.][]{hudson2002generating}
but with several supporting a range of different models
of sequence
evolution~\citep[e.g.][]{mailund2005coasim,excoffier2011fastsimcoal,
virgoulay2021gspace}.

\begin{figure}
    \includegraphics{figures/unmutated_tree}
    \includegraphics{figures/mutated_tree}
    \caption{
        \label{fig-mutated-trees}
        Visualization of the separation between ancestry and mutation simulation:
        \texttt{sim\_ancestry} produces ancestry
        (\textbf{left;} shown are two trees along a 1Kb chunk of genome
        relating 200 samples),
        and \texttt{sim\_mutations} adds mutations
        (\textbf{right;} fourteen nucleotide mutations have been added).
        Graphics produced with \tskit's \texttt{draw\_svg} method.
        \plrcomment{TODO: fix up the axis labels, if we like this figure?}
    }
\end{figure}
